---
title: "R Basics"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
date: "2024-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Course text

The course book is https://epirhandbook.com/

# Installing R and Rstudio
Installing R and RStudio https://posit.co/download/rstudio-desktop/

The RStudio environment https://epirhandbook.com/en/r-basics.html#rstudio


# Functions and arguments
Functions and arguments https://epirhandbook.com/en/r-basics.html#functions


# Packages
Install and loadhttps://epirhandbook.com/en/r-basics.html#packages

Install using the function `install.packages()`
Remember to the put the package name in quotation marks e.g.

`install.packages("tidyverse")`

`install.packages("rio")`

Load packages using `library()`

`library(tidyverse)`

`library(rio)`

No quotation marks required


```{r include=FALSE}
library(tidyverse)
library(rio)
```


# Scripts
https://epirhandbook.com/en/r-basics.html#scripts

File ending in `.R`

You can comment out text in scripts using #
```{r}
# Comment
```

You can use the keyboard shortcut `Ctrl + shift + c` to comment

A more useful script is R markdown `.Rmd` (see below)

# Objects
https://epirhandbook.com/en/r-basics.html#objects

Use the 'assignment operator' `<-` to create an object

```{r}
x <- 49

x2 <- sqrt(x)
```


# Importing data
https://epirhandbook.com/en/import-and-export.html

A dummy linelist can be found here https://github.com/appliedepi/epirhandbook_eng/raw/master/data/case_linelists/linelist_cleaned.xlsx
You can download this and save on your machine in a data folder

```{r}
# linelist <- import("data/linelist_cleaned.xlsx")
```

You can also import directly from the web

```{r}
linelist <- import("https://github.com/appliedepi/epirhandbook_eng/raw/master/data/case_linelists/linelist_cleaned.xlsx")
```


# Working directory
https://epirhandbook.com/en/r-basics.html#working-directory

getwd()

# R projects
https://epirhandbook.com/en/r-projects.html#r-projects

You can create projects from git repositories (see below)

# File paths

Note slash direction

- Windows - `C:\Users\ashley.sharp\Desktop\R projects\rtraining\data`

- R - `C:/Users/ashley.sharp/Desktop/R projects/rtraining/data`

- Absolute paths - avoid

`linelist <- import("C:/Users/ashley.sharp/Desktop/R projects/rtraining/data/linelist_cleaned.xlsx")`

- Relative paths - better

`linelist <- import("data/linelist_cleaned.xlsx")`


# Object classes

Numeric

```{r}
x <- 49
class(x)
```

Character/string (use quotation marks)

```{r}
string <- "abc"
class(string)
class("data/linelist_cleaned.xlsx")
```

Datetime (POSIXct (calendar time) and POSIXlt (local time))

```{r}
now <- Sys.time()
class(now)
```

Data frame
```{r}
class(linelist)
```


# Reviewing data

```{r}
library(skimr)

skim(linelist)
```




# R markdown
<https://epirhandbook.com/en/reports-with-r-markdown.html>

File ending in .Rmd

A more flexible kind of script that you can use to make automated documents
 <https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>

Rmd files have several features:

## Headings

-   Bullets

*Italic*

**Bold**

<!-- comment out this text -->
<!-- Comment out text with Ctrl + Shift + c -->

Code chunks

```{r}
# This is a code chunk
```

Outputs

```{r}
print("this is an output")
```


# Git and github
https://epirhandbook.com/en/version-control-and-collaboration-with-git-and-github.html


Install git https://git-scm.com/downloads

Git repository for course materials url: https://github.com/ashleysharp/rtraining


Create a project from git repository:
File -> new project -> create project from version control -> enter url and choose location on your machine

Get latest code from git repository
Git -> Pull



