---
title: "RMarkdown"
output: html_document
date: "2024-09-05"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Reports and R markdown

A useful summary can be found in the Epi R handbook
<https://epirhandbook.com/en/reports-with-r-markdown.html>

<https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>

## Key points

File ending in .Rmd

The working directory

The working directory of a markdown file is wherever the Rmd file itself
is saved. For instance, if the R project is within \~/Documents/projectX
and the Rmd file itself is in a subfolder
\~/Documents/projectX/markdownfiles/markdown.Rmd, the code
read.csv(“data.csv”) within the markdown will look for a csv file in the
markdownfiles folder, and not the root project folder where scripts
within projects would normally automatically look.

There are three basic components to an Rmd file: YAML, Markdown text,
and R code chunks.

### YAML metadata

Referred to as the ‘YAML metadata’ or just ‘YAML’, this is at the top of
the R Markdown document. This section of the script will tell your Rmd
file what type of output to produce, formatting preferences, and other
metadata such as document title, author, and date.

### Markdown text

Titles and headings A hash symbol in a text portion of a R Markdown
script creates a heading. This is different than in a chunk of R code in
the script, in which a hash symbol is a mechanism to
comment/annotate/de-activate, as in a normal R script.

# First-level heading / title

## Second level heading

### Third-level heading

Bullets and numbering Use asterisks (\*) to created a bullets list.
Finish the previous sentence, enter two spaces, Enter/Return twice, and
then start your bullets. Include a space between the asterisk and your
bullet text. After each bullet enter two spaces and then Enter/Return.
Sub-bullets work the same way but are indented. Numbers work the same
way but instead of an asterisk, write 1), 2), etc. Below is how your R
Markdown script text might look.

Here are my bullets (there are two spaces after this colon):

-   Bullet 1 (followed by two spaces and Enter/Return)\
-   Bullet 2 (followed by two spaces and Enter/Return)
    -   Sub-bullet 1 (followed by two spaces and Enter/Return)\
    -   Sub-bullet 2 (followed by two spaces and Enter/Return)

*Italic*

**Bold**

Comment out text
<!-- You can “comment out” R Markdown text just as you can use the “#” to comment out a line of R code in an R chunk. Simply highlight the text and press Ctrl+Shift+c (Cmd+Shift+c for Mac). The text will be surrounded by arrows and turn green. It will not appear in your output. -->

### Code chunks

```{r}
# This is a code chunk
```

Outputs

```{r}
print("this is an output")
```

Sections of the script that are dedicated to running R code are called
“chunks”. This is where you may load packages, import data, and perform
the actual data management and visualisation.

Each chunk is opened with a line that starts with three back-ticks, and
curly brackets that contain parameters for the chunk ({ }). The chunk
ends with three more back-ticks.

You can create a new chunk by typing it out yourself, by using the
keyboard shortcut “Ctrl + Alt + i” (or Cmd + Shift + r in Mac), or by
clicking the green ‘insert a new code chunk’ icon at the top of your
script editor.

### In text r code

Today's date is `r Sys.Date()`

### Templates

Useful templates are available in the ukhsacharts package
<https://gitlab.phe.gov.uk/fs-data-science-leads/ukhsacharts>

-   HTML UKHSA template
-   HTML UKHSA slidy template
-   PowerPoint UKHSA template
-   Word UKHSA template

Once the package is installed you should be able to access these
templates (you may need to restart R Studio). To do this, click:

File New File R Markdown... From Template These templates should be
listed with ukhsacharts in curly brackets on the right hand side.
